\ChUnnumbered{Introduction}{Intro}

\p The \acrfull{hlsl} is the GPU programming language originally provided in
conjunction with the \gls{dx} runtime. Over many years its use has expanded to
cover every major rendering API across all major development platforms.

\p \acrshort{hlsl} draws heavy inspiration originally from \gls{isoC} and later
from \gls{isoCPP} with additions specific to graphics and parallel computation
programming. The language is also influenced to a lesser degree by other popular
graphics and parallel programming languages.

\p \acrshort{hlsl} has two reference implementations which this specification
draws heavily from. The original reference implementation \acrfull{fxc} has been
in use since \gls{dx} 9. The more recent reference implementation \acrfull{dxc}
has been the primary shader compiler since \gls{dx} 12.

\p In writing this specification bias is leaned toward the language
behavior of \acrshort{dxc} rather than the behavior of \acrshort{fxc}, although
that can vary by context.

\p In very rare instances this spec will be aspirational, and may diverge from
both reference implementation behaviors. This will only be done in instances
where there is an intent to alter implementation behavior in the future. Since
this document and the implementations are living sources, one or the other may
be ahead in different regards at any point in time. The authors will see to
provide editorial notes in such cases to clarify the intent of the specification
and the state of the implementations.

\p This document is a \textit{working draft}, and has not been adopted by the
Ecma General Assembly. As such, this document is subject to change as the
specification process continues.

% TODO: Delete the paragraph above and replace it with the text below:
% \p This Ecma Standard was developed by Technical Committee 57 and was adopted
% by the General Assembly of <month> <year>.
