\Ch{Resource Types}{Resources}

\p Resource types are intangible types representing memory or configuration with
an external interface.

\p These take the form of typed buffers, raw buffers, textures, constant buffers
and samplers.

\p Typed buffer, raw buffer, and Texture resources may be read-only or
read-write as denoted by the type of the resource, where the names of read-write
resource types are prefixed with \texttt{RW}.

\p Global resource declarations are \textit{bound} to device memory managed by
an application through a runtime implementation. A resource declaration that is
not bound is said to be \textit{unbound}. Operations on unbound resources are
evaluated as no-ops.

\p An implementation may impose restrictions on how local resource objects may
be used (\ref{AnnexB.LocalResources}).

\Sec{Optional Features}{Resources.Optional}

\p \textit{Dynamic resources} is an optional feature which allows resource
objects to be initialized by indexing directly into a heap of descriptors. A
conforming implementation must either support all interfaces related to dynamic
resources or none, but may not partially support them.

\p The \texttt{\_\_DynamicResourceHandle} type in the code examples in this
clause is a placeholder for the implementation-specific type returned by
subscript indexing into the \texttt{ResourceDescriptorHeap}
(\ref{Resources.Binding.Dynamic}).

\Sec{Special Semantics}{Resources.Special}

\Sub{Resource Initialization}{Resources.Special.Initialization}

\p A resource declaration in a global scope are called \textit{global
resource declaration} and represents a binding for external data into a
program. Global resource declarations are implicitly initialized in an
implementation-defined way.

\p Use of an uninitialized resource is undefined.

\Sec{Typed Buffers}{Resources.TypedBuffers}

\p The typed buffer class template represents a one-dimensional resource
containing an array of a single given type. Its contents are indexed by typed
access to each element. The memory backing a resource need not formatted in the
same type as the resource's element type. If required, typed buffer elements are
converted upon load in an implementation-defined way.

\p There are two typed buffer class templates \texttt{Buffer} and
\texttt{RWBuffer}, which respectively represent read-only and read-write typed
buffer objects.

\begin{HLSL}
template <typename T = float4>
requires hlsl::is_vector<T>::value || hlsl::is_arithmetic<T>::value
class Buffer {
  Buffer();
  Buffer(const Buffer &buf);
  Buffer &operator=(Buffer &buf);
  Buffer(const __DynamicResourceHandle &handle);

  void GetDimensions(out uint width) const;

  // Element access.
  T operator[](int index) const;
  T Load(int index) const;
  T Load(int index, out uint status) const;
};


template <typename T>
requires hlsl::is_vector<T>::value || hlsl::is_arithmetic<T>::value
class RWBuffer {
  RWBuffer();
  RWBuffer(RWBuffer buf);
  RWBuffer &operator=(RWBuffer &buf);
  Buffer(const __DynamicResourceHandle &handle);

  void GetDimensions(out uint width) const;

  // Element access.
  T operator[](int index) const;
  device T &operator[](int n) const;
  T Load(int index) const;
  T Load(int index, out uint status) const;
};
\end{HLSL}

\p The element type of a typed buffer, \texttt{T}, must be an arithmetic type
(\ref{Basic.Types.Arithmetic}) or vector of arithmetic containing no more than 4
elements, and the total size of the element (\texttt{sizeof(T)}) must not exceed
16-bytes in size.

\p The \texttt{Buffer} type's template argument \texttt{T} has a default value
of \texttt{float4}.

\Sub{Typed Buffer Constructors}{Resources.TypedBuffers.Ctrs}

\begin{HLSL}
Buffer();
RWBuffer();
\end{HLSL}

\begin{itemize}
  \item \textit{Effects:} Constructs an uninitialized resource handle.
\end{itemize}

\begin{HLSL}
Buffer(const Buffer &buf);
RWBuffer(const RWBuffer &buf);
\end{HLSL}

\begin{itemize}
  \item \textit{Effects:} Initializes a resource to refer to the same resource
  as the input argument \texttt{buf}.
\end{itemize}

\begin{HLSL}
Buffer(const __DynamicResourceHandle &handle);
\end{HLSL}

\begin{itemize}
  \item \textit{Feature:} Dependent on dynamic resources (\ref{Resources.Binding.Dynamic}).
  \item \textit{Effects:} Initializes a resource to refer to the dynamic
  resource handle provided as the input argument \texttt{handle}.
\end{itemize}

\Sub{Typed Buffer Copying and Assignment}{Resources.TypedBuffers.Copy}

\begin{HLSL}
Buffer &operator=(Buffer &buf);
RWBuffer &operator=(RWBuffer &buf);
\end{HLSL}

\begin{itemize}
  \item \textit{Effects:} Initializes a resource to refer to the same resource
  as the input argument \texttt{buf}.
  \item \textit{Returns:} \texttt{this}.
\end{itemize}

\Sub{Typed Buffer Inspecting Functions}{Resources.TypedBuffers.Inspect}

\begin{HLSL}
void GetDimensions(out uint width) const;
\end{HLSL}

\begin{itemize}
  \item \textit{Effects:} Writes the total number of elements contained in the
  typed buffer to the \texttt{width} argument.
\end{itemize}

\begin{HLSL}
T Load(uint index) const;
\end{HLSL}

\begin{itemize}
  \item \textit{Requires:} \texttt{index} \textless the number of elements in the
  typed buffer.
  \item \textit{Returns:} The element in the typed buffer at the index specified
  by \texttt{index}.
\end{itemize}

\begin{HLSL}
T Load(int index, out uint status) const;
\end{HLSL}

\begin{itemize}
  \item \textit{Requires:} \texttt{index} \textless the number of elements in the
  typed buffer.
  \item \textit{Returns:} The element in the typed buffer at the index specified
  by \texttt{index}.
  \item \textit{Effects:} Initializes the \texttt{status} argument to an
  implementation-defined status signal (\ref{Resources.Mapcheck}).
  \item \textit{Remarks:} The \texttt{status} parameter returns an
  implementation-defined value that indicates if the loaded value came from
  fully mapped parts of the resource. This parameter must be passed to the
  built-in function \texttt{CheckAccessFullyMapped} (\ref{Resources.Mapcheck})
  in order to interpret it.
\end{itemize}

\Sub{Typed Buffer Operators}{Resources.TypedBuffers.Operators}

\begin{HLSL}
T operator[](uint index) const;
\end{HLSL}

\begin{itemize}
  \item \textit{Requires:} \texttt{index} \textless the number of elements in the
  typed buffer.
  \item \textit{Returns:} The element in the typed buffer at the index specified
  by \texttt{index}.
\end{itemize}

\begin{HLSL}
device T &operator[](int index);
\end{HLSL}

\begin{itemize}
  \item \textit{Preconditions:} \texttt{index} \textless the number of elements in the
  typed buffer.
  \item \textit{Returns:} An lvalue-reference to the element in the typed buffer
  at the index specified by \texttt{index}.
  \item \textit{Remarks:} Writing to typed buffer elements overwrite the entire
  element, even if only part of the data is written. A partial write to a typed
  buffer is a load-insert-store operation.
\end{itemize}

\begin{note}
  Accurately modeling the memory operations for writable typed buffers would
  require a proxy-object interface since the address written to isn't actually
  the device memory since it may be converted or otherwise normalized by the
  implementation.
\end{note}

\Sec{CheckAccessFullyMapped}{Resources.Mapcheck}

\p The implementation-defined status value returned by texture and buffer load
methods is interpreted by a built-in function:

\begin{HLSL}
bool CheckAccessFullyMapped(uint status);
\end{HLSL}

\begin{itemize}
  \item \textit{Preconditions:} \texttt{status} is an initialized value from a
  resource \texttt{Load} function's \texttt{status} output parameter.
  \item \textit{Returns:} \texttt{true} if the value was accessed from memory
  that was fully mapped by the runtime implementation; otherwise returns
  \texttt{false}.
\end{itemize}
